---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

read data from "activity.csv" to a data.frame, "alldata".

```{r}
alldata<-read.csv("activity.csv")
alldata$interval<-sapply(alldata$interval,function(x) paste(paste(rep("0",4-nchar(as.character(x))),sep="",collapse = ''),as.character(x),sep=""))
alldata$date <- as.Date(alldata$date, format = "%Y-%m-%d")
```
copy all the non-NA observation to a data.frame, "cleandata".
```{r}
cleandata <- alldata[!is.na(alldata$steps),]
```

## What is mean total number of steps taken per day?

Calculate the averaged total steps per day and put them in "data_steps".

```{r}
date_steps<-aggregate(cleandata$steps,list(date=cleandata$date),sum)
colnames(date_steps)[2]<-"total_steps"
hist(date_steps$total_steps,main="total steps per day",xlab="")
```

```{r}
mean(date_steps$total_steps)
```
The mean of total steps per day are 11766.19.

```{r}
median(date_steps$total_steps)

```
The median of total steps per day are `r median(date_steps$total_steps)`.

## What is the average daily activity pattern?

```{r}
interval_steps<-aggregate(cleandata$steps,list(inte=cleandata$interval),mean)
plot(strptime(interval_steps$inte,"%H%M"),interval_steps$x,type = "l",xlab="interval",ylab="mean steps")
```

Calculate the interval which has the maximum mean steps:
```{r}
idx<-which(interval_steps$x==max(interval_steps$x))
interval_steps$inte[idx]
```

The mean steps has its maximum at interval 8:35.

## Imputing missing values

Calculate the total number of missing values in the dataset:
```{r}
sum(is.na(alldata$steps))
```
The total number of missing values is 2304.

Fill the NA slots with the mean value of the steps at the same interval of other days.

First we create a function that takes values of steps from the corresponding time period if the value of steps is NA.
```{r}
rmna <- function (x,y) {
  if(is.na(x)) {
    interval_steps[which(interval_steps$inte==(y)),2]
  } else {x}
}
```
Now mapply the function to create a data set "alldata2" with all NA value filled.
```{r}
alldata2<-alldata
alldata2$steps<-mapply(rmna,alldata$steps,alldata$interval)
```
plot the histogram again with alldata2.
```{r}
date_steps2<-aggregate(alldata2$steps,list(date=alldata2$date),sum)
colnames(date_steps2)[2]<-"total_steps"
hist(date_steps2$total_steps,main="total steps per day",xlab="")
```

```{r}
mean(date_steps2$total_steps)
```
The mean of total steps per day are the same as before, 11766.19.

```{r}
median(date_steps2$total_steps)

```
The median of total steps per day becomes 11766.19, the same as the mean value.

## Are there differences in activity patterns between weekdays and weekends?
